Evaluation of classical correlation functions from 2/3D images on CPU and GPU architectures: introducing CorrelationFunctions.jl

### 2022-10-07 paper-correlations.tex
Add many formulas to look like a smart guy

\section{Verifying our algorithms}
\label{sec:verification}
To verify our package we use random datasets of overlapping disks (the number of
dimensions $n = 2$) or balls ($n = 3$) with radii $R$ and centers placed in
points generated by Poisson point process with a parameter $\lambda$. For these
datasets correlation functions can be given in closed-form expressions for one-,
two- and three-dimensional cases. For the sake of testing we obtain a collection
of disks (resp. balls) in a square (resp. cube) with the side $s$ using the
following algorithm:
\begin{algorithmic}[1]
  \label{testdata}
  \Procedure{testdata}{$s, n, R, \lambda$}
    \State $nballs \gets poisson(\lambda s^n)$ \Comment{$poisson(\lambda)$
      generates random numbers with Poisson distribution with parameter
      $\lambda$.}
    \State Generate $nballs$ random points which are uniformly distributes in
    a cube ($n = 3$) or square ($n = 2$) with a side $s$
    \State Draw balls (disks) of radius $R$ in those points. Store the result in
    an array $output$.
    \State \textbf{return} $output$
  \EndProcedure
\end{algorithmic}

In two-dimensional case we test functions $L_2(r)$, $S_2(r)$, $F_{ss}(r)$ and
$F_{sv}(r)$ using parameters $R = 100$, $\lambda = 4\cdot10^{-5}$ and
$s = 5000$. The correlation functions are calculated in two axial directions
with periodic boundary conditions and the result is averaged across the
directions. Instead of taking one realization of Poisson process we take 15
realizations and average results over them. The result can be seen on
\cref{fig:2d}.

For pore size function $P(r)$ and chord length function $p(\bm{r})$ a slightly
different approach is required. We take a single realization of Poisson point
process with the same parameters $R$ and $\lambda$ in a square with the side
$s = 20000$ (resp. $s = 12000$) and calculate histograms for Pore size
(resp. Chord length) function. Chord length function is calculated in two axial
directions. The theoretical values are integrated over bin ranges and compared
with the number of elements in those bins. The results are presented on
\cref{fig:pscl}.

In three-dimensional case the approach for testing is still the same but with
different parameters to speed up calculations. We choose $R = 20$,
$\lambda = 3\cdot10^{-5}$ and $s = 500$. The correlation functions are
calculated in three axial directions with periodic boundary conditions and the
result is also averaged across the directions as in two-dimensional
case. We take six realizations of Poisson process for averaging the results
which are presented on \cref{fig:3d}.

Pore size and chord length functions are calculated using the side of a cube
$s = 700$. The results are presented on \cref{fig:pscl}.

The results for three-dimensional case are seemingly worse than for
two-dimensional case. This can be explained by a reduced number of realizations
of Poisson process and smaller value of $s$.

>>>>

\section{Verifying our algorithms}
\label{sec:verification}
Our algorithms are verified by comparing their results with known closed-form
expressions for correlation functions for some sort of input data. One kind of
well-studied data is overlapping n-balls with fixed radius and centers generated
by Poisson point process. Here we consider two-dimensional and three-dimensional
cases.

\subsection{Two-dimensional case}
Consider an infinite set of disks with radii $R$ and centers generated by
Poisson point process with parameter $\lambda$. \textcolor{red}{Insert picture}.

Two-point function in this case is: \cite{Torquato_book}
\begin{equation*}
  S_2(r) = e^{-2\lambda \left\{
  \begin{array}{ll}
    \pi R^2 & \quad r > 2R \\
    \pi R^2 + r\sqrt{A}/4 - BR^2 & \quad \text{otherwise}
  \end{array}
  \right.}
\end{equation*}
Here and later $A = 4R^2 - r^2$ and $B = \arccos \frac{r}{2R}$.

An expression for $F_{ss}$ is
\begin{equation*}
  F_{ss}(r) = 4S_2(r) \left\{
  \begin{array}{ll}
    (\pi\lambda R)^2 & \quad r > 2R \\
    \frac{AR^2\lambda^2 r(B^2 - 2\pi B + \pi^2) + \sqrt{A}R^2\lambda}{Ar} & \quad \text{otherwise}
  \end{array}
  \right.
\end{equation*}

Surface-void function $F_{sv}$ becomes
\begin{equation*}
  F_{sv}(r) = 2R\lambda S_2(r) \left\{
  \begin{array}{ll}
    \pi & \quad r > 2R \\
    \pi-B & \quad \text{otherwise}
  \end{array}
  \right.
\end{equation*}

Two-point function $L_2$:
\begin{equation*}
  L_2(r) = e^{-\lambda(\pi R^2 + 2rR)}
\end{equation*}

Chord length function $p$:
\begin{equation*}
  p(r) = 2\lambda R e^{-2\lambda rR}
\end{equation*}

And finally pore size function $P$:
\begin{equation*}
  P(r) = 2\lambda\pi (r+R) e^{-\lambda\pi(r^2 + 2rR)}
\end{equation*}

\textcolor{red}{Add results}

\subsection{Three-dimensional case}
Like in the previous case, consider an infinite set of balls with radii $R$ and
centers generated by Poisson point process with parameter $\lambda$. Closed-form
expressions for correlations functions are also known for this case.

Two-point function in this case is:
\begin{equation*}
  S_2(r) = e^{-\lambda\pi \left\{
  \begin{array}{ll}
    \frac{8}{3} R^3 & \quad r > 2R \\
    \frac{4}{3} R^3 + rR^2 - \frac{1}{12}r^3 & \quad \text{otherwise}
  \end{array}
  \right.}
\end{equation*}

Let $\eta = \frac{4}{3}\lambda \pi R^3$. Then an expression for $F_{ss}$ is
\begin{equation*}
  F_{ss}(r) = S_2(r) \left\{
  \begin{array}{ll}
    \frac{9\eta^2}{R^2} & \quad r > 2R \\
    \frac{9\eta^2}{R^2}(\frac{1}{2}+\frac{r}{4R})^2 + \frac{3\eta}{2rR} & \quad \text{otherwise}
  \end{array}
  \right.
\end{equation*}

Surface-void function $F_{sv}$ is:
\begin{equation*}
  F_{sv}(r) = \frac{3\eta}{R} S_2(r) \left\{
  \begin{array}{ll}
    1 & \quad r > 2R \\
    \frac{1}{2} + \frac{r}{4R} & \quad \text{otherwise}
  \end{array}
  \right.
\end{equation*}

Two-point function $L_2$:
\begin{equation*}
  L_2(r) = e^{-\lambda\pi (\frac{4}{3}R^3 + rR^2)}
\end{equation*}

Chord length function $p$:
\begin{equation*}
  p(r) = \pi\lambda R^2 e^{-\pi\lambda rR^2}
\end{equation*}

And finally pore size function $P$:
\begin{equation*}
  P(r) = 4\pi\lambda(r+R)^2 e^{-\frac{4}{3}\pi\lambda (r^3 + 3r^2R + 3rR^2)}
\end{equation*}

\textcolor{red}{Add results}
